What We Eval in the Shadows

### 2021-08-09 paper/main.tex
Add more examples

\begin{lstlisting}
rARS <- function(n, formula, min=-Inf, max=Inf, sp) {
    ...
    p <- function(x) { eval(parse(text=formula)) }
    ...
}
\end{lstlisting}

@@

\begin{lstlisting}
.makeCppMethods <- function(methods, cppclassinfo, env) {
    cppMethods <- names(cppclassinfo@methods)
    newMethods <- names(methods)
    for(what in cppMethods[! cppMethods %in% newMethods])
        methods[[what]] <- eval(substitute(
                  function(...) .CppObject$WHAT(...), list(WHAT = as.name(what))),
                                env)
    methods
}
\end{lstlisting}

@@

\begin{lstlisting}
NVL <- function(...) {
    for(e in eval(substitute(alist(...))))
        x <- eval(e, parent.frame())
        if(!is.null(x)) break
    }
    x
}
\end{lstlisting}
The expression \code{eval(substitute(alist(...)))} is used to extract the
unevaluated text of the arguments passed to \code{NVL}.
\code{substitute(alist(...))} returns an expression representing a call to the
\code{alist} function with \code{...} replaced by the arguments supplied to
\code{NVL}. Evaluating this expression returns a \code{list} whose arguments are
the unevaluated argument expressions which are evaluated sequentially for their
result.

@@

\begin{lstlisting}
`%||%` <- function (lhs, rhs) {
    lres <- withVisible(eval(lhs, envir = parent.frame()))
    if (is.null(lres$value)) {
        eval(rhs, envir = parent.frame())
    } else {
        if (lres$visible) { lres$value}
        else { invisible(lres$value) }
   }
}
\end{lstlisting}

@@

\begin{lstlisting}
see_if <- function(..., env = parent.frame(), msg = NULL) {
    asserts <- eval(substitute(alist(...)))
    for (assertion in asserts) {
      res <- tryCatch({
        eval(assertion, env)
      }, assertError = function(e) {
        structure(FALSE, msg = e$message)
      })
    ...
}

\end{lstlisting}

@@

\begin{lstlisting}
do_call <- function(f, ..., .args = list(), .env = parent.frame()) {
  f <- substitute(f)
  call <- make_call(f, ..., .args)
  eval(call, .env)
}
\end{lstlisting}

@@

\begin{lstlisting}
library(glue)
greeting <- "Hello"
glue('{greeting} World!')
#> Hello World!
\end{lstlisting}

@@

\emph{Avoid Boilerplate.}

For example, the \code{PopED::get\_rse} uses it to assign a number of variables
in the current scope: \code{eval(parse(text=paste(default_args[[i]]), "<- ",
  i))}. There are also more sophisticated cases. The \code{plyr::allocate\_column}
function uses eval for a deferred assignment to fill missing values in a column.
Package overture repeatedly evaluates R expressions capturing any assignments
that happen in it as samples of Markov chains.

\begin{lstlisting}
scalaEnsure <- function() {
    ...
    env <- parent.env(environment())
    eval(parse(text=paste0('unlockBinding("s", env)')))
    assign("s", s, envir=env)
    lockBinding("s", env)
}
\end{lstlisting}

\subsection{Domain Specific Language}

@@

>>>>

\begin{figure}[h]
\begin{lstlisting}
rARS <- function(n, formula, min=-Inf, max=Inf, sp) {
    ...
    p <- function(x) { eval(parse(text=formula)) }
    ...
}
\end{lstlisting}
  \caption{AdapSamp::rARS}
\end{figure}

@@

\begin{figure}[h]
\begin{lstlisting}
.makeCppMethods <- function(methods, cppclassinfo, env) {
    cppMethods <- names(cppclassinfo@methods)
    newMethods <- names(methods)
    for(what in cppMethods[! cppMethods %in% newMethods])
        methods[[what]] <- eval(substitute(
                  function(...) .CppObject$WHAT(...), list(WHAT = as.name(what))),
                                env)
    methods
}
\end{lstlisting}
  \caption{Rcpp::.makeCppMethods}
\end{figure}

@@

\begin{figure}[h]
\begin{lstlisting}
NVL <- function(...) {
    for(e in eval(substitute(alist(...))))
        x <- eval(e, parent.frame())
        if(!is.null(x)) break
    }
    x
}
\end{lstlisting}
  \caption{statnet.common::NVL}
\end{figure}

The expression \code{eval(substitute(alist(...)))} is used to extract the
unevaluated text of the arguments passed to \code{NVL}.
\code{substitute(alist(...))} returns an expression representing a call to the
\code{alist} function with \code{...} replaced by the arguments supplied to
\code{NVL}. Evaluating this expression returns a \code{list} whose arguments are
the unevaluated argument expressions which are evaluated sequentially for their
result.

@@

\begin{figure}[h]
\begin{lstlisting}
`%||%` <- function (lhs, rhs) {
    lres <- withVisible(eval(lhs, envir = parent.frame()))
    if (is.null(lres$value)) {
        eval(rhs, envir = parent.frame())
    } else {
        if (lres$visible) { lres$value}
        else { invisible(lres$value) }
   }
}
\end{lstlisting}
  \caption{igraph::\%||\%}
\end{figure}

@@

\begin{figure}[h]
\begin{lstlisting}
see_if <- function(..., env = parent.frame(), msg = NULL) {
    asserts <- eval(substitute(alist(...)))
    for (assertion in asserts) {
      res <- tryCatch({
        eval(assertion, env)
      }, assertError = function(e) {
        structure(FALSE, msg = e$message)
      })
    ...
}
\end{lstlisting}
  \caption{assertthat::see\_if}
\end{figure}

@@

\begin{figure}[h]
\begin{lstlisting}
do_call <- function(f, ..., .args = list(), .env = parent.frame()) {
  f <- substitute(f)
  call <- make_call(f, ..., .args)
  eval(call, .env)
}
\end{lstlisting}
\caption{igraph::do\_call}
\end{figure}

@@

\begin{figure}[h]
\begin{lstlisting}
library(glue)
greeting <- "Hello"
glue('{greeting} World!')
#> Hello World!
\end{lstlisting}
  \caption{glue::glue}
\end{figure}

@@

@@

\emph{Easier Maintenance.} The combination of \c{match.call} and \eval is
frequently used to make it easier to maintain functions with similar interfaces.
This pattern is very commonly employed by modeling functions. Consider the
example below of the \code{survfit.formula} function from package
\code{survival}.

\begin{figure}[h]
\begin{lstlisting}
survfit.formula <- function(formula, data, ...) {
  Call <- match.call()
  indx <- match(c('formula', 'data', ...), names(Call), nomatch=0)
  temp <- Call[c(1, indx)]
  temp[[1L]] <- quote(stats::model.frame)
  mf <- eval.parent(temp)
  ...
}
\end{lstlisting}
  \caption{survival::survfit.formula}
\end{figure}

This function uses \code{match.call} to extract the expression with which it is
called. Then, it extracts the indices of relevant arguments using the
\code{match} function. Next, it subsets the \code{Call} variable to extract
elements at those indices along with the first element which is the function
name. The function name is then replaced with \code{stats::model.frame}.
Finally, this expression is evaluated. By using \code{...}, the function accepts
arbitrary arguments. It filters the arguments of interest using \code{match}
function. \todo{explain more}

\emph{Remove Boilerplate.} For example, the \code{PopED::get\_rse} uses it to
assign a number of variables in the current scope:
\code{eval(parse(text=paste(default_args[[i]]), "<- ", i))}. There are also more
sophisticated cases. The \code{plyr::allocate\_column} function uses eval for a
deferred assignment to fill missing values in a column. Package overture
repeatedly evaluates R expressions capturing any assignments that happen in it
as samples of Markov chains.

\begin{figure}[h]
\begin{lstlisting}
scalaEnsure <- function() {
    ...
    env <- parent.env(environment())
    eval(parse(text=paste0('unlockBinding("s", env)')))
    assign("s", s, envir=env)
    lockBinding("s", env)
}
\end{lstlisting}
\caption{aibd::scalaEnsure}
\end{figure}

\emph{Lookup and indexing.}
